\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2021-12-11 sections/methods.tex

  \subsection{Consensus method}
  \vspace{-5mm}

  The consensus algorithm was designed to increase the number of true positives reported at the end of the network inference step. For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter \(p\). The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter \(p\). Here, the \(N_{i}\) each have the same set of nodes and only differ by the value of the association inferred between every pair.

  \subsection{Algorithm 1: Simple voting}
  \label{sec:org34a8cdc}

@@

  \begin{equation}
  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor p * x \rfloor
  \label{eqn:simple-voting}
  \end{equation}
  where, \(x\) is the number of network inference methods used in the consensus.

@@

  \subsection{Algorithm 2: Scaled-sum method}
  \label{sec:orgb0ddb1f}

  This algorithm performs a consensus based on sum of all edges reported between a node-pair. First, the network generated by each network inference method is re-scaled so that \(\max(\mid N_{i} \mid) = 1\) (Equation \ref{eqn:scaled-sum-rescaling}).

@@

  The re-scaling is performed because the associations reported by the direct association metrics are not necessarily correlations and are not directly comparable to the associations from the correlation-based methods due to their lower weights. This problem can also be alleviated by weighing each network inference method differently. Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum})

@@

  \subsection*{Network Variability}
  \vspace{-5mm}
 In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules and parameters, we organized multiple networks into a single mathematical structure that we could use for linear regression.
 In particular, we transformed the adjacency matrix of each co-occurrence network into a vector.
 We then merged the networks generated from all possible combinations of tools into a table (N, see below) in which each column represents one network.

@@

  \subsubsection*{p-value merging}
  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
  \begin{equation*}
    \begin{aligned}
        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
        \Psi &\sim \chi^2_{2k}
    \end{aligned}
  \end{equation*}

>>>>

  \subsubsection*{p-value merging (can be moved to the supplementary section)}
  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
  \begin{equation*}
    \begin{aligned}
        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
        \Psi &\sim \chi^2_{2k}
    \end{aligned}
  \end{equation*}

 \subsection*{NOTATION}

Note this section defines the notation for the network comparison component of our pipeline, which assumes that all networks to be compared have the same number of nodes

q = number of nodes

w = number of different networks from different approaches (previously x?)

$N^i$ = network k (currently Ni)

$N^i_{a,b}$ = edge (a,b) in network i

It useful to define a "stretched" version of each adjacency matrix $N^i$ into a column vector, where all columns are stacked onto each other into an $q^2$ long vector:

$\bar{N}^i_{j}$ , $j=1:q^2$

  \subsection*{Consensus method}
  \vspace{-5mm}

  The consensus algorithm was designed to increase the number of true positives reported at the end of the network inference step. For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter \(p\). The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter (p). Here, the \(N_{i}\) each have the same set of nodes and only differ by the value of the association inferred between every pair.

  \subsection*{Algorithm 1: Simple voting}

@@

 For each edge (i,j), we count number $M_{i,j}$ of networks in which that edge exists, i.e.:

 \begin{equation}
 \begin{aligned}
 M_{i,j} &= | \left{ a | N^{(a)}_{i,j}>0 \right} | \\
 M_{i,j} &\geq \lfloor \theta * w \rfloor
 \end{aligned}
  \label{eqn:simple-voting}
  \end{equation}

  \begin{equation}
  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor \theta * w \rfloor
  \label{eqn:simple-voting}
  \end{equation}

   \begin{equation}
   \sum_k N^k_{i,j} \geq \lfloor \theta * w \rfloor
  \label{eqn:simple-voting2}
  \end{equation}

  where, \(w\) is the number of network inference methods used in the consensus.

 where, \theta is the user-defined threshold....

@@

  \subsubsection*{Algorithm 2: Scaled-sum method}

  This algorithm performs a consensus based on sum of all edges reported between a node-pair. First, the network generated by each network inference method is re-scaled so that (\max(\mid N_{i} \mid) = 1) (Equation \ref{eqn:scaled-sum-rescaling}).

\begin{equation}
  NS^{a}_{i,j} = \frac{N^{(a)}_{i,j}}{\max(\mid N_{i} \mid)}
  \label{eqn:scaled-sum-rescaling}
  \end{equation}

@@

  The re-scaling is performed because the associations reported by the direct association metrics are not necessarily correlations and are not directly comparable to the associations from the correlation-based methods due to their lower weights. This problem can also be alleviated by weighing each network inference method differently. Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum}).

@@

  \subsubsection*{Network Variability}
  \vspace{-5mm}
 In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules and parameters, we organized multiple networks into a single mathematical structure that we could use for linear regression.
 In particular, we transformed the adjacency matrix of each co-occurrence network into a vector.
 We then merged the networks generated from all possible combinations of tools into a table (N, see below) in which each column represents one network.

@@

  The possible correlations between the edges of the network lead to inaccurate estimates of the variance if a GLM were used to directly model the relationships between the links and steps in the workflow. Therefore, we transformed the network matrix $N$ using a Principal Component Analysis into the $N_{reduced}$ matrix. The dimensions of the $N_{reduced}$ matrix are be denoted by $(d \times n)$ where d is the number of principal components and n is the number of network in the $N$ matrix.
